---
title: "Frontloading GO"
author: "Samuel Gurr"
date: "2024-02-21"
output: pdf_document
---


### Set working directory
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      cache = TRUE)

knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/Airradians_CellularMolecular_OA/RAnalysis/Output/Transcriptomics/WGCNA/challenge/Frontloaded_Activated/gProfiler")
#knitr::opts_knit$set(root.dir = "C:/Users/samuel.gurr/Documents/Github_repositories/Airradians_CellularMolecular_OA/RAnalysis")
```

### Load libraries
```{r load_libraries, include = TRUE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(tidyverse)
library(ggpubr)
```


### Load data: 

**Data:** Data are from (gprofiler)[https://biit.cs.ut.ee/gprofiler/gost] online! All genes with LOC gene names for Cvirgninca (blast hits to Airraidans) were copied and pasted into g profiler to get GO Enrichment. 

**Objective:** Build line plots for th elog10 p values of highlighted terms output from gprofiler

**What are the HIGHLIGHTED terms from p:Profler??** 
g:Profiler highlights certain GO terms to *reduce redundancy and improve the interpretability of functional enrichment* results. This highlighting is achieved through a novel two-stage algorithm implemented in g:Profiler The algorithm starts with the term having the smallest adjusted p-value and excludes its child and ancestor terms from further searches. It then recalculates the hypergeometric p-value for remaining terms, excluding genes already seen in the previous significant term

```{r gprofiler outputs}

# frontloaded - criteia subset from module pink + magenta results 
gprofiler.Frontloaded        <- read.csv(file = "frontloaded/Frontloaded_GO_master.csv",header = T)

# costly activated - criteia subset from module red and brown results
gprofiler.CostlyActivation   <- read.csv(file = "costly_activation/CostlyActivation_GO_master.csv",header = T)

```


## Segment plots

```{r plot frontloaded}

Segplot.Frontloaded <- gprofiler.Frontloaded %>% 
                            # filter the data for target rows
                            dplyr::filter(highlighted %in% 'TRUE' & !source %in% 'GO:CC') %>% 
                            dplyr::mutate(log10_adjusted_p_value = abs(log10(adjusted_p_value)),
                                          source = gsub('.*GO:','',source)) %>% # remove GO: %>% 
                              # now ggplot
                              ggplot(aes(x=term_name, y= log10_adjusted_p_value, size = intersection_size)) + 
                                geom_segment(aes(x=term_name, 
                                                 xend=term_name, 
                                                 y=1, 
                                                 yend=log10_adjusted_p_value, 
                                                 color=source,
                                                 size = 3)) + 
                                geom_point(aes(size = intersection_size), shape =21,  fill = "white") +
                                coord_flip() +
                                theme(panel.grid.minor.y = element_blank(),
                                      panel.grid.major.y = element_blank(),
                                      legend.position="bottom") +
                                xlab("") +
                                ylab("") +
                                ggtitle("WGCNA Frontloaded gProfiler") + 
                                theme_bw() + #Set background color 
                                theme(panel.border = element_blank(), # Set border
                                      panel.grid.major = element_blank(), #Set major gridlines
                                      panel.grid.minor = element_blank(), #Set minor gridlines
                                      axis.line = element_line(colour = "black"), #Set axes color
                                      plot.background=element_blank()) + #Set the plot background #set title attributes
                                facet_wrap(source ~., scales="free_y", ncol= 1, strip.position="right", shrink = T)


pdf(paste0("frontloaded/Frontloaded_SegPlot.pdf"), width=5, height=5)
plot(Segplot.Frontloaded)
dev.off()


```

```{r plot costly activation}

Segplot.CostlyActivation <- gprofiler.CostlyActivation %>% 
                            # filter the data for target rows
                            dplyr::filter(highlighted %in% 'TRUE' & !source %in% 'GO:CC') %>% 
                            dplyr::mutate(log10_adjusted_p_value = abs(log10(adjusted_p_value)),
                                          source = gsub('.*GO:','',source)) %>% # remove GO: %>% 
                              # now ggplot
                              ggplot(aes(x=term_name, y= log10_adjusted_p_value, size = intersection_size)) + 
                                geom_segment(aes(x=term_name, 
                                                 xend=term_name, 
                                                 y=1, 
                                                 yend=log10_adjusted_p_value, 
                                                 color=source,
                                                 size = 3)) + 
                                geom_point(aes(size = intersection_size), shape =21,  fill = "white") +
                                coord_flip() +
                                theme(panel.grid.minor.y = element_blank(),
                                      panel.grid.major.y = element_blank(),
                                      legend.position="bottom") +
                                xlab("") +
                                ylab("") +
                                ggtitle("WGCNA Costly Activation gProfiler") + 
                                theme_bw() + #Set background color 
                                theme(panel.border = element_blank(), # Set border
                                      panel.grid.major = element_blank(), #Set major gridlines
                                      panel.grid.minor = element_blank(), #Set minor gridlines
                                      axis.line = element_line(colour = "black"), #Set axes color
                                      plot.background=element_blank()) + #Set the plot background #set title attributes
                                facet_wrap(source ~., scales="free_y", ncol= 1, strip.position="right", shrink = T)


pdf(paste0("costly_activation/CostlyActivation_SegPlot.pdf"), width=5, height=5)
plot(Segplot.CostlyActivation)
dev.off()


```