---
title: "Hemolymph_analysis"
author: "Samuel Gurr"
date: "2023-05-09"
output: html_notebook
---

-   Last updates: May 5/9/2023

# OA Reexposure Challenge

### May 2023: F2 Adult Bay scallops exposed to a two-week full-reciprocal pCO2 challenge

### Objective: plot and analyze flow cytometry data for (a) SYBR green and propidium iodide - cell viability (b) MitoSox gree - mitochondrial ROS (c) JC-10 - mitochondrial membrane potential corrected for CCCP

## Load Libraries

```{r startup, include=FALSE}

library(ggplot2)
library(dplyr)
library(tidyverse)
library(car)
library(multcompView)
library(agricolae)
library(Rmisc)
library(lmerTest)
library(pander)
library(performance)
library(Rmisc)
library(tidyr)
library(reshape2)
library(rcompanion)
library(kableExtra)
library(ggpubr)
library(tidyr)
library(nlme)
library(car)
library(rcompanion) # to run the Schrier -Ray-Hare non parametric 2 way 
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/Airradians_Reexposure_OA/RAnalysis") # sets the working directory for the entire R markdown file - no need to reload the wd
```

## Load flow cytometry data

**NOTE**: the statistics option ion C6 PLus software was used to compile the data (i.e. mean FL1) - the following loaded data file was assembled manually to include the sample metadata (i.e. data, treatments, etc.) as well as these flow cytometry data

```{r load data}
data  <-read.csv(file="Data/FlowCytometry/Hemolymph_data_raw.csv", header = TRUE)
head(data)
```

## Plot data 

### Cell viability (Live vs. dead)

* SYBRgreen + propidium iodide, gated using the FL2 to reveal true peaks (PE-H) 

* data comprised of percent live and dead

```{r plot live dead}

Day1_PeacLive_RxnNorm <- data %>%
  # edit data to suit out needs
  dplyr::filter(Date %in% 20230502) %>% # filter for desired date of rhte experiment
  dplyr::mutate(SYBR_PI_percent_alive = as.numeric(sub("%", "", SYBR_PI_percent_alive))) %>% # the % sign is included making this a character, omit the percent and make numeric!
  dplyr::mutate(SYBR_PI_percent_alive = as.numeric(sub(",", "", SYBR_PI_percent_alive))) %>% # the % sign is included making this a character, omit the percent and make numeric!
  na.omit() %>%  # omitts sample ID 64 that did not have sufficient hemolymph for this assay 
  #dplyr::select(c('pCO2_exposure', 'pCO2_history','SYBR_PI_count_alive')) %>%  # select desired columns
  group_by(pCO2_exposure, pCO2_history) %>% # group by columns for treatment
  dplyr::summarise( # summarise to aquire the mean and SE for plotting
    PercLive_mean = mean(SYBR_PI_percent_alive), # mean
    PercLive_sd = sd(SYBR_PI_percent_alive), # sd
    n = n(), # count
    PercLive_se = PercLive_sd / sqrt(n)) %>% # SE
  # plot it
  ggplot(aes(x=pCO2_exposure, y=PercLive_mean, group=pCO2_history)) + 
    geom_line(aes(group = factor(pCO2_history)), size = 0.5) +  # connect a line between variables
    geom_point(aes(shape=pCO2_history, fill=pCO2_history), size = 3) + 
    scale_shape_manual(values=c(16, 17, 4)) + # filled circle, filled triangle, and X 
    scale_fill_manual(values=c('white','white', 'black')) + # fill cicle white, triangle white, and jsut black for the X
    geom_errorbar(aes(ymin=(PercLive_mean)-(PercLive_se), 
                      ymax=(PercLive_mean)+(PercLive_se)), 
                  width=.2,position=position_dodge(.1)) +
    # geom_jitter() +
    theme_classic() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    ggtitle("Percent alive cells - 24 hours") +
    labs(y= "Percent alive cells (%)", x = "pCO2 Exposure") +
    facet_wrap(~pCO2_history)
Day1_PeacLive_RxnNorm



Day1_PercDead_RxnNorm <- data %>%
  # edit data to suit out needs
  dplyr::filter(Date %in% 20230502) %>% # filter for desired date of rhte experiment
  dplyr::mutate(SYBR_PI_percent_dead = as.numeric(sub("%", "", SYBR_PI_percent_dead))) %>% # the % sign is included making this a character, omit the percent and make numeric!
  dplyr::mutate(SYBR_PI_percent_dead = as.numeric(sub(",", "", SYBR_PI_percent_dead))) %>% # the % sign is included making this a character, omit the percent and make numeric!
  na.omit() %>%  # omitts sample ID 64 that did not have sufficient hemolymph for this assay 
  #dplyr::select(c('pCO2_exposure', 'pCO2_history','SYBR_PI_count_dead')) %>%  # select desired columns
  group_by(pCO2_exposure, pCO2_history) %>% # group by columns for treatment
  dplyr::summarise( # summarise to aquire the mean and SE for plotting
    PercDead_mean = mean(SYBR_PI_percent_dead), # mean
    PercDead_sd = sd(SYBR_PI_percent_dead), # sd
    n = n(), # count
    PercDead_se = PercDead_sd / sqrt(n)) %>% # SE
  # plot it
  ggplot(aes(x=pCO2_exposure, y=PercDead_mean, group=pCO2_history)) + 
    geom_line(aes(group = factor(pCO2_history)), size = 0.5) +  # connect a line between variables
    geom_point(aes(shape=pCO2_history, fill=pCO2_history), size = 3) + 
    scale_shape_manual(values=c(16, 17, 4)) + # filled circle, filled triangle, and X 
    scale_fill_manual(values=c('white','white', 'black')) + # fill cicle white, triangle white, and jsut black for the X
    geom_errorbar(aes(ymin=(PercDead_mean)-(PercDead_se), 
                      ymax=(PercDead_mean)+(PercDead_se)), 
                  width=.2,position=position_dodge(.1)) +
    # geom_jitter() +
    theme_classic() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    ggtitle("Percent dead cells - 24 hours") +
    labs(y= "Percent dead cells (%)", x = "pCO2 Exposure") +
    facet_wrap(~pCO2_history)
Day1_PercDead_RxnNorm



```


### Mitochondrial reactive oxygen species 

* MitoSox Green, gated using the FL1 (FITC-H) and FL1 vs. FL2 to truncate cells (FITC-H v. PE-H)

* data is the Mean FL1 

```{r plot Mitochondrial ROS}

Day1_MitoROS_RxnNorm <- data %>%
  # edit data to suit out needs
  dplyr::filter(Date %in% 20230502) %>% # filter for desired date of rhte experiment
  na.omit() %>%  # omitts sample ID 64 that did not have sufficient hemolymph for this assay 
  #dplyr::select(c('pCO2_exposure', 'pCO2_history','SYBR_PI_count_alive')) %>%  # select desired columns
  group_by(pCO2_exposure, pCO2_history) %>% # group by columns for treatment
  dplyr::summarise( # summarise to aquire the mean and SE for plotting
    MitROS_mean = mean(MitoSoxGreen_Mean_FL1), # mean
    MitROS_sd = sd(MitoSoxGreen_Mean_FL1), # sd
    n = n(), # count
    MitROS_se = MitROS_sd / sqrt(n)) %>% # SE
  # plot it
  ggplot(aes(x=pCO2_exposure, y=MitROS_mean, group=pCO2_history)) + 
    geom_line(aes(group = factor(pCO2_history)), size = 0.5) +  # connect a line between variables
    geom_point(aes(shape=pCO2_history, fill=pCO2_history), size = 3) + 
    scale_shape_manual(values=c(16, 17, 4)) + # filled circle, filled triangle, and X 
    scale_fill_manual(values=c('white','white', 'black')) + # fill cicle white, triangle white, and jsut black for the X
    geom_errorbar(aes(ymin=(MitROS_mean)-(MitROS_se), 
                      ymax=(MitROS_mean)+(MitROS_se)), 
                  width=.2,position=position_dodge(.1)) +
    # geom_jitter() +
    theme_classic() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    ggtitle("Mitochondrial ROS - 24 hours") +
    labs(y= "A.U.", x = "pCO2 Exposure") +
    facet_wrap(~pCO2_history)
Day1_MitoROS_RxnNorm

```


### Mitochondrial membrane potential

* JC-10 (compensated for CCCP), gated using the FL1 (FITC-H) and FL1 vs. FL2 to truncate cells (FITC-H v. PE-H)

* data is the ratio of 'red to gree' as the Jaggregate (positive membrane potential) : Monomer (low memrane potential) or the FL 2 : FL 1 

```{r plot Mitochondrial membrane potential}

data$JC10_FL2_FL1_Ratio <- ( (as.numeric(sub(",", "", data$JC10_FL2_Jaggregate))) / 
                             (as.numeric(sub(",", "", data$JC10_FL1_monomer))) )


Day1_MitoMemPot_RxnNorm <- data %>%
  # edit data to suit out needs
  dplyr::filter(Date %in% 20230502) %>% # filter for desired date of rhte experiment
  na.omit() %>%  # omitts sample ID 64 that did not have sufficient hemolymph for this assay 
  #dplyr::select(c('pCO2_exposure', 'pCO2_history','SYBR_PI_count_alive')) %>%  # select desired columns
  group_by(pCO2_exposure, pCO2_history) %>% # group by columns for treatment
  dplyr::summarise( # summarise to aquire the mean and SE for plotting
    JC10_mean = mean(JC10_FL2_FL1_Ratio), # mean
    JC10_sd = sd(JC10_FL2_FL1_Ratio), # sd
    n = n(), # count
    JC10_se = JC10_sd / sqrt(n)) %>% # SE
  # plot it
  ggplot(aes(x=pCO2_exposure, y=JC10_mean, group=pCO2_history)) + 
    geom_line(aes(group = factor(pCO2_history)), size = 0.5) +  # connect a line between variables
    geom_point(aes(shape=pCO2_history, fill=pCO2_history), size = 3) + 
    scale_shape_manual(values=c(16, 17, 4)) + # filled circle, filled triangle, and X 
    scale_fill_manual(values=c('white','white', 'black')) + # fill cicle white, triangle white, and jsut black for the X
    geom_errorbar(aes(ymin=(JC10_mean)-(JC10_se), 
                      ymax=(JC10_mean)+(JC10_se)), 
                  width=.2,position=position_dodge(.1)) +
    # geom_jitter() +
    theme_classic() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
    ggtitle("Mitochondrial membrane potential - 24 hours") +
    labs(y= "A.U.", x = "pCO2 Exposure") +
    facet_wrap(~pCO2_history)
Day1_MitoMemPot_RxnNorm


```

```{r output plots by date} 


pdf(paste0("C:/Users/samjg/Documents/Github_repositories/Airradians_Reexposure_OA/RAnalysis/Output/FlowCytometry/Day1_Hemolymph_RxnNorm.pdf"), height = 12)
ggarrange(Day1_PeacLive_RxnNorm, Day1_PercDead_RxnNorm, Day1_MitoROS_RxnNorm, Day1_MitoMemPot_RxnNorm, ncol=1)
dev.off()


```
